---
title: "Animal Model example"
author: "August Arnstad"
date: "`r format(Sys.time(), '%d.%m.%Y')`"
output:
  pdf_document:
    toc: yes
    toc_depth: '2'
  html_document:
    toc: yes
    toc_float: yes
    toc_depth: 2
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(formatR)
showsol <- FALSE
library(knitr)
library(devtools)
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 68), 
                      tidy = TRUE, 
                      warning = FALSE, 
                      error = FALSE, 
                      message = FALSE, 
                      echo = TRUE, 
                      fig.width=7, 
                      fig.height=5, 
                      fig.align="center")

library(remotes)
library(devtools)
if(!require(devtools)) install.packages("devtools")
devtools::install_github("AugustArnstad/BayesianVariableImportance")
library(BayesianVariableImportance)
library(ggplot2)
library(INLA)
library(microbenchmark)
library(Matrix)
```

## This code contains the models fitted and samples drawn, without needing to access the data:

```{r}
# Access the stored models and samples if you just want to see the results. 
# The code with the original data used in the thesis is found further down below.
# These are samples with the CCD integration, but grid integration is also in the folder

setwd("/your/desired/path/BayesianVariableImportance/Vignettes/Animal model example")

model_mass <- readRDS("Models and samples/model_mass_ccd_24.05.rds")
samples_mass <- readRDS("Models and samples/samples_mass_ccd_24.05.rds")

model_wing <- readRDS("Models and samples/model_wing_ccd_24.05.rds")
samples_wing <- readRDS("Models and samples/samples_wing_ccd_24.05.rds")

model_tarsus <- readRDS("Models and samples/model_tarsus_ccd_24.05.rds")
samples_tarsus <- readRDS("Models and samples/samples_tarsus_ccd_24.05.rds")
```

```{r}
plots_mass <- BayesianVariableImportance::plot_samples(samples_mass)
plots_wing <- BayesianVariableImportance::plot_samples(samples_wing)
plots_tarsus <- BayesianVariableImportance::plot_samples(samples_tarsus)

```

```{r, echo = FALSE, eval = FALSE}
plots_mass$fixed_effects
plots_mass$random_effects
plots_mass$heritability
plots_mass$R2

plots_wing$fixed_effects
plots_wing$random_effects
plots_wing$heritability
plots_wing$R2

plots_tarsus$fixed_effects
plots_tarsus$random_effects
plots_tarsus$heritability
plots_tarsus$R2
```




# Custom plots heritability for use in thesis.
# In general, I think I should pick a colour palette that is the same for the whole thesis.
```{r}
str(samples_mass)

silva_mass <- 0.300
silva_wing <- 0.388
silva_tarsus <- 0.415

muff_mass <- 0.288
muff_wing <- 0.344

additive_genetic <- "IDC2"

CI_mass <- quantile(samples_mass$scaled_random_samples$IDC2, c(0.025, 0.975))


# Dynamically specify the column name in aes() using rlang's sym() and !! for tidy evaluation
heritability_mass <- ggplot(samples_mass$scaled_random_samples, aes(x = !!sym(additive_genetic))) +
  geom_histogram(aes(y = ..density..), fill = "#C6CDF7", alpha = 0.7, bins = 40, color = "black") +
  geom_density(color = "#E6C6DF", adjust = 1.5, linewidth=1.5) +
  annotate("point", x = mean(samples_mass$scaled_random_samples$IDC2), y = 0, color = "#E6C6DF", size = 4) +
  annotate("point", x = median(samples_mass$scaled_random_samples$IDC2), y = 0, color = "grey", size = 4) +
  annotate("point", x = mean(silva_mass), y = 0, color = "#FFD700", size = 4) +
  annotate("point", x = mean(muff_mass), y = 0, color = "#32CD32", size = 4) +
  geom_vline(aes(xintercept = CI_mass[1]), color = "#E6C6DF", linewidth = 1, linetype = "dashed") +
  geom_vline(aes(xintercept = CI_mass[2]), color = "#E6C6DF", linewidth = 1, linetype = "dashed") +
  labs(#title = paste("Heritability of mass"),
       x = "Samples of heritability",
       y = "Frequency") +
  theme_minimal() +
  theme(legend.position = "none",
    axis.title.x = element_text(size = 24),
    axis.title.y = element_text(size = 24),
    axis.text.x = element_text(size = 24),
    axis.text.y = element_text(size = 24)
    ) 


heritability_mass

mean(samples_mass$scaled_random_samples$IDC2)
median(samples_mass$scaled_random_samples$IDC2)
```


```{r}
str(samples_wing)

CI_wing <- quantile(samples_wing$scaled_random_samples$IDC2, c(0.025, 0.975))

# Dynamically specify the column name in aes() using rlang's sym() and !! for tidy evaluation
heritability_wing <- ggplot(samples_wing$scaled_random_samples, aes(x = !!sym(additive_genetic))) +
  geom_histogram(aes(y = ..density..), fill = "#C6CDF7", alpha = 0.7, bins = 40, color = "black") +
  geom_density(color = "#E6C6DF", adjust = 1.5, linewidth=1.5) +
  annotate("point", x = mean(samples_wing$scaled_random_samples$IDC2), y = 0, color = "#E6C6DF", size = 4) +
  annotate("point", x = median(samples_wing$scaled_random_samples$IDC2), y = 0, color = "grey", size = 4) +
  annotate("point", x = mean(silva_wing), y = 0, color = "#FFD700", size = 4) +
  annotate("point", x = mean(muff_wing), y = 0, color = "#32CD32", size = 4) +
  geom_vline(aes(xintercept = CI_wing[1]), color = "#E6C6DF", linewidth = 1, linetype = "dashed") +
  geom_vline(aes(xintercept = CI_wing[2]), color = "#E6C6DF", linewidth = 1, linetype = "dashed") +
  labs(#title = paste("Heritability of wing"), 
       x = "Samples of heritability", 
       y = "Frequency") +
  theme_minimal() +
  theme(legend.position = "none",
    axis.title.x = element_text(size = 24),
    axis.title.y = element_text(size = 24),
    axis.text.x = element_text(size = 24),
    axis.text.y = element_text(size = 24)
    ) 


heritability_wing

mean(samples_wing$scaled_random_samples$IDC2)
median(samples_wing$scaled_random_samples$IDC2)

quantile(samples_wing$scaled_random_samples$IDC2, c(0.025, 0.975))
```


```{r}
str(samples_tarsus)

CI_tarsus <- quantile(samples_tarsus$scaled_random_samples$IDC2, c(0.025, 0.975))

# Dynamically specify the column name in aes() using rlang's sym() and !! for tidy evaluation
heritability_tarsus <- ggplot(samples_tarsus$scaled_random_samples, aes(x = !!sym(additive_genetic))) +
  geom_histogram(aes(y = ..density..), fill = "#C6CDF7", alpha = 0.7, bins = 40, color = "black") +
  geom_density(color = "#E6C6DF", adjust = 1.5, linewidth=1.5) +
  annotate("point", x = mean(samples_tarsus$scaled_random_samples$IDC2), y = 0, color = "#E6C6DF", size = 4) +
  annotate("point", x = median(samples_tarsus$scaled_random_samples$IDC2), y = 0, color = "grey", size = 4) +
  annotate("point", x = mean(silva_tarsus), y = 0, color = "#FFD700", size = 4) +
  geom_vline(aes(xintercept = CI_tarsus[1]), color = "#E6C6DF", linewidth = 1, linetype = "dashed") +
  geom_vline(aes(xintercept = CI_tarsus[2]), color = "#E6C6DF", linewidth = 1, linetype = "dashed") +
  labs(#title = paste("Heritability of tarsus"), 
       x = "Samples of heritability", 
       y = "Frequency") +
  theme_minimal() +
  theme(legend.position = "none",
    axis.title.x = element_text(size = 24),
    axis.title.y = element_text(size = 24),
    axis.text.x = element_text(size = 24),
    axis.text.y = element_text(size = 24)
    ) 


heritability_tarsus

mean(samples_tarsus$scaled_random_samples$IDC2)
median(samples_tarsus$scaled_random_samples$IDC2)
quantile(samples_tarsus$scaled_random_samples$IDC2, c(0.025, 0.975))
sd(samples_tarsus$scaled_random_samples$IDC2)#/sqrt(length(samples_tarsus$scaled_random_samples$IDC2))

```


## The below code is made with the data d.morph used in the thesis. A pedigree structure is inialized and the corresponding inverse relatedness matrix is called Cmatrix. Note that the thesis used both integration strategies, but the user can make the choice of prefered integration strategy.

```{r, echo = FALSE, eval = FALSE}
data("d.morph")
data("Cmatrix")

str(d.morph)
isSymmetric(Cmatrix)
```


```{r}
formula.mass = mass ~ sex + FGRM + month + age + outer + other +
  f(hatchyear,model="iid",hyper=list(
    prec=list(initial=log(1), prior="pc.prec",param=c(1,0.05))
  ))+
  f(IDC,model="iid",hyper=list(
    prec=list(initial=log(1), prior="pc.prec",param=c(1,0.05))
  )) +
  f(IDC2,values=1:3116,model="generic0",
    Cmatrix=Cmatrix,
    constr = TRUE,
    hyper=list(
      # The priors are relevant, need to discuss
      prec=list(initial=log(0.5), prior="pc.prec",param=c(sqrt(2),0.05))
    ))

formula.tarsus = tarsus ~ sex + FGRM + month + age + outer + other +
  f(hatchyear,model="iid",hyper=list(
    prec=list(initial=log(1), prior="pc.prec",param=c(1,0.05))
  ))+
  f(IDC,model="iid",hyper=list(
    prec=list(initial=log(1), prior="pc.prec",param=c(1,0.05))
  )) +
  f(IDC2,values=1:3116,model="generic0",
    Cmatrix=Cmatrix,
    constr = TRUE,
    hyper=list(
      # The priors are relevant, need to discuss
      prec=list(initial=log(0.5), prior="pc.prec",param=c(sqrt(2),0.05))
    ))

formula.wing = wing ~ sex + FGRM + month + age + outer + other +
  f(hatchyear,model="iid",hyper=list(
    prec=list(initial=log(1), prior="pc.prec",param=c(1,0.05))
  ))+
  f(IDC,model="iid",hyper=list(
    prec=list(initial=log(1), prior="pc.prec",param=c(1,0.05))
  )) +
  f(IDC2,values=1:3116,model="generic0",
    Cmatrix=Cmatrix,
    constr = TRUE,
    hyper=list(
      # The priors are relevant, need to discuss
      prec=list(initial=log(0.5), prior="pc.prec",param=c(sqrt(2),0.05))
    ))
```

```{r, echo = FALSE, eval = FALSE}
set.seed(1)
prior_mass = list(theta = list(initial=log(0.5),  prior="pc.prec",param=c(sqrt(2),0.05)))

timing_result_mass <- microbenchmark(
  {
    model_mass <- BayesianVariableImportance::perform_inla_analysis(d.morph, formula.mass, "gaussian", int_strat="ccd", priors = prior_mass)

    samples_mass <- BayesianVariableImportance::sample_posterior_gaussian(model_mass, formula.mass, d.morph, n_samp=100, additive_param = "IDC2")
  },
  times = 1  # Number of times to run the chunk for better accuracy
)

print(timing_result_mass)

plots_mass <- BayesianVariableImportance::plot_samples(samples_mass)
plots_mass$fixed_effects
BayesianVariableImportance::plot_posteriors_and_heritability(model_mass)
```

```{r, echo = FALSE, eval = FALSE}
plots_mass$fixed_effects
plots_mass$random_effects
plots_mass$heritability
plots_mass$R2
```

```{r}
set.seed(1)
prior_wing = list(theta = list(initial=log(0.5),  prior="pc.prec",param=c(sqrt(2),0.05)))
#prior_wing = list(theta = list(initial=log(1),  prior="pc.prec",param=c(sqrt(2),0.05)))

timing_result_wing <- microbenchmark(
  {
    model_wing <- BayesianVariableImportance::perform_inla_analysis(d.morph, formula.wing, "gaussian", int_strat="ccd", priors = prior_wing)

    samples_wing <- BayesianVariableImportance::sample_posterior_gaussian(model_wing, formula.wing, d.morph, n_samp=10000, additive_param = "IDC2")
  },
  times = 1  # Number of times to run the chunk for better accuracy
)

print(timing_result_wing)


plots_wing <- BayesianVariableImportance::plot_samples(samples_wing)
```

```{r}
plots_wing$fixed_effects
plots_wing$random_effects
plots_wing$heritability
plots_wing$R2
```

```{r}
set.seed(1)
prior_tarsus = list(theta = list(initial=log(0.5),  prior="pc.prec",param=c(sqrt(2),0.05)))

timing_result_tarsus <- microbenchmark(
  {
      model_tarsus <- BayesianVariableImportance::perform_inla_analysis(d.morph, formula.tarsus, "gaussian", int_strat="ccd", priors = prior_tarsus)

      samples_tarsus <- BayesianVariableImportance::sample_posterior_gaussian(model_tarsus, formula.tarsus, d.morph, n_samp=10000, additive_param = "IDC2")
  },
  times = 1  # Number of times to run the chunk for better accuracy
)

print(timing_result_tarsus)


plots_tarsus <- BayesianVariableImportance::plot_samples(samples_tarsus)
```

```{r}
plots_tarsus$fixed_effects
plots_tarsus$random_effects
plots_tarsus$heritability
plots_tarsus$R2

dim(d.morph)
```


```{r}
# Saving the entire loaded_sim_no$successes object as an RDS file
setwd("/your/desired/path/Animal model results")

saveRDS(model_mass, "model_mass_XX.XX.rds")
saveRDS(samples_mass, "samples_mass_XX.XX.rds")

saveRDS(model_wing, "model_wing_XX.XX.rds")
saveRDS(samples_wing, "samples_wing_XX.XX.rds")

saveRDS(model_tarsus, "model_tarsus_XX.XX.rds")
saveRDS(samples_tarsus, "samples_tarsus_XX.XX.rds")
```


